Constraining Quiescent Supermassive Black Hole Environments.

### 2014-09-01 ms.tex
Filling in paragraphs

\newcommand{\Menc}{M_{\rm enc}}
\newcommand{\rhostar}{\rho_*}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\Mseight}{M_{\star,8}}
\newcommand{\Mbh}[1][]{M_{\bullet#1}}
\newcommand{\MbhNorm}{\frac{\Mbh}{10^8 \Msun}}

@@

\newcommand{\vwO}{v_{w,0}}
\newcommand{\kewO}{\frac{\vwO^2}{2}}
\newcommand{\x}{\frac{r_s}{\rsoi}}
\newcommand{\vwNorm}{\frac{\vwO}{\sigsoi}}
\newcommand{\vwNormB}{\frac{\vwO}{500 \, {\rm km/s}}}
\newcommand{\vwNormC}{\frac{\vwO}{500 \, {\rm km/s}}}

@@

%In this study we perform hydrodynamic modeling to solve for the steady-state gas density, temperature and velocity profiles near the SMBH sphere of influence, for a sample of nearby, quiescent galaxies. We a adopt the model used by \cite{Quataert:2004a} for the galactic center: gas is supplied by stellar winds, which collide, shock heat, and thermalize.  
We take the same approach. To the best of our knowledge a systematic study of the parameter space of central galaxy properties. Thus, we compute central density profiles for a broad sample of galaxies from \citealt{WangMerritt:2004a}--henceforward \citetalias{WangMerritt:2004a}.  This allows us to relate the central stellar properties (e.g. the slope of the stellar density) to the central gas properties.

@@

%%AG--some discussion of alternate parameterizations.  My intuition is that for our purposes the parameterization should not matter. But some discussion of alternate possibilities could prevent Alistar Graham from yelling at us.

@@

%{\bf NCS: here is some added detail about the role of stellar mass functions in the model}

% if we consider different star formation scenarios (for example, continuous star formation) or different IMFs, it will change.  Once these free parameters are specified, however, we can answer an important question: do young stellar populations increase or decrease the SMBH feeding rate $\dot{M}$?  Clearly, $\eta(t)$ is larger for young stellar populations, but these stars also have high wind velocities that diminish the stagnation radius.  Crudely approximating $v_{\rm w}=200~{\rm km~s}$ for $m_{\rm max} < 10M_{\odot}$ and $v_{\rm w}=1000~{\rm km~s}$ for $m_{\rm max} > 10M_{\odot}$ (motivated by the transition from dust-driven wind loss on the AGB to line-driven wind loss on the main sequence), we can employ the relation $\dot{M} \propto \eta(t) r_{\rm s}^{2-\Gamma}\propto \eta(t) v_{\rm w}^{-4+2\Gamma}$ (where $\rho_* \propto r^{-\Gamma}$) to determine the impact of stellar ``youth'' on SMBH feeding rates.
%
%\begin{figure}
%\includegraphics[width=\columnwidth]{NickPlot.eps}
%\caption{\label{NickPlot} SMBH feeding rates $\dot{M}=\eta(t) \times M_*(r_{\rm s})$, in arbitrary units.  The green, blue, and black curves are for galaxies with $\Gamma$ values of $0.1$, $0.5$, and $0.9$, respectively.  Solid curves represent impulsive-mode star formation, while dotted curves represent continuous-mode star formation.}
%\end{figure}
%
%%In Fig. \ref{NickPlot} we plot $\dot{M}$, in arbitrary units, as a function of time, for three different stellar density profiles $\Gamma = \{1.1, 1.5, 1.9\}$ (which are normalized to have the same mass at an influence radius $r_{\rm soi}=10~{\rm pc}$ around a $10^7M_\odot$ SMBH).  We parametrize the wind velocity as 
%\begin{equation}
%\frac{v_{\rm w}}{\rm km~s^{-1}}=600-400\tanh\left( \frac{t-10^{7.5}~{\rm yr}}{10^{7}~{\rm yr}}\right).
%\end{equation}
%The ``impulsive burst'' mode of star formation produces large ($\sim 10$) differences between the three $\dot{M}$ curves at early times, when $r_{\rm s}$ is small, but smaller ($\sim 3$) differences at late times, when $r_{\rm s}$ is large.  We also plot, as dotted curves, a simple model for the ``continuous'' mode of star formation, where mass loss is calculated as $\bar{\eta} = \int\eta(t)dt/t_{\rm h} \approx 4$ and an average energy injection in the wind is calculated as $\bar{v_{\rm w}^2}=\int\eta(t)v_{\rm w}^2(t)dt/\bar{\eta} \approx (800~{\rm km~s}^{-1})^2$.  Interestingly, the continuous mode of star formation produces small differences from late-time $\dot{M}$ seen in cusp galaxies with impulsive star formation; however, continuous mode star formation decreases late-time $\dot{M}$ by an order of magnitude relative to impulsive star formation in core galaxies.

@@

\begin{align}
&\rs=\frac{7}{2}\frac{G \Mbh}{n \vw^2}\\
&\simeq 6 \, \pc \left(\MbhNorm\right) \left(\vwNormB\right)^{-2}
\label{eq:stag_simple}
\end{align}

@@

\[
\dot{M}
\begin{cases}
   1.5 \time 10^{-3}  \left(\MbhNorm\right)^{1.9} \left(\vwNormB\right)^{-4} \Msun \, \pyear& \text{for core galaxies} \\
   3.3 \times 10^{-3}   & \text{for cusp galaxies}     
\end{cases}
\label{eq:cooling}
\]

@@

%%AG-Apparently even in the low heating case, it is still possible to an outflow (from the analytic results above the gas density must be quite steep in this case). However, at low heating we will probably become subject to a cooling instablity.
%%AG-I believe that in the case of core galaxy, the negative root of the quadratic should be taken...

@@

The stagnation radius scales linearly with $\Mbh$ (see Equation~\ref{eq:stag_simple}), while the $\rIa$ is roughly constant. For small $\Mbh$, $\rs$ is well inside $\rIa$ and SNes will not contribute to the $\vwO$.  Above $\sim10^8 \Msun$ SNes will represent an important contribution to the overall heating. 
%Thus, for $\Mbh=10^6 \Msun$, we may have $\vwO\simeq 100$ km/s with heating mostly coming from main sequence winds. On the other hand, for $\Mbh \gsim 10^8 \Msun$, we may have $\vwO\simeq 500$ km/s, with the main contribution coming from type Ia SNe. These two regimes are illustrated in Figure~\ref{fig:vw_eff}.
%\end{enumerate}

@@

%AG-I am not comfortable saying type Ia SNe's don't contribute to heating for smaller mass BHs. If the cooling time for the SNe is long then it would be wrong to ignore the SNe contribution. The difference would be that the SNe heating would just be less steady.
%AG-Thornton and http://arxiv.org/pdf/1402.6695v2.pdf...
\end{enumerate}

@@

The gas should radiatively cool as well--either from metal line cooling for $T\lsim 2\times 10^7$ K, or through free-free emission for $T\gsim 2\times 10^7$ K.  However, if the cooling rate is much smaller than the rate of energy injection then radiative cooling may be safely neglected.

@@

>>>>

\newcommand{\Menc}{M_{\rm enc}}
\newcommand{\rhostar}{\rho_*}
\newcommand{\Mstar}{M_{\star}}
\newcommand{\Mseight}{M_{\star,8}}
\newcommand{\Mbh}[1][]{M_{\bullet#1}}
\newcommand{\Mbheight}{M_{\bullet,8}}
\newcommand{\MbhNorm}{\frac{\Mbh}{10^8 \Msun}}

@@

\newcommand{\vwO}{v_{w,0}}
\newcommand{\kewO}{\frac{\vwO^2}{2}}
\newcommand{\x}{\frac{r_s}{\rsoi}}
\newcommand{\vwNorm}{\frac{\vwO}{\sigsoi}}
\newcommand{\vwOFH}{v_{w,0,500}}
\newcommand{\vwOFHexp}{\frac{\vwO}{500 \, {\rm km/s}}}

@@

% In this study we perform hydrodynamic modeling to solve for the
% steady-state gas density, temperature and velocity profiles near the
% SMBH sphere of influence, for a sample of nearby, quiescent
% galaxies. We a adopt the model used by \cite{Quataert:2004a} for the
% galactic center: gas is supplied by stellar winds, which collide,
% shock heat, and thermalize.
We take the same approach. To the best of our knowledge a systematic
study of the parameter space of central galaxy properties. Thus, we
compute central density profiles for a broad sample of galaxies from
\citealt{WangMerritt:2004a}--henceforward
\citetalias{WangMerritt:2004a}.  This allows us to relate the central
stellar properties (e.g. the slope of the stellar density) to the
central gas properties.

@@

%% AG--some discussion of alternate parameterizations.  My intuition
%% is that for our purposes the parameterization should not
%% matter. But some discussion of alternate possibilities could
%% prevent Alistar Graham from yelling at us.

@@

% {\bf NCS: here is some added detail about the role of stellar mass
% functions in the model}

% if we consider different star formation scenarios (for example,
% continuous star formation) or different IMFs, it will change.  Once
% these free parameters are specified, however, we can answer an
% important question: do young stellar populations increase or
% decrease the SMBH feeding rate $\dot{M}$?  Clearly, $\eta(t)$ is
% larger for young stellar populations, but these stars also have high
% wind velocities that diminish the stagnation radius.  Crudely
% approximating $v_{\rm w}=200~{\rm km~s}$ for $m_{\rm max} <
% 10M_{\odot}$ and $v_{\rm w}=1000~{\rm km~s}$ for $m_{\rm max} >
% 10M_{\odot}$ (motivated by the transition from dust-driven wind loss
% on the AGB to line-driven wind loss on the main sequence), we can
% employ the relation $\dot{M} \propto \eta(t) r_{\rm
% s}^{2-\Gamma}\propto \eta(t) v_{\rm w}^{-4+2\Gamma}$ (where $\rho_*
% \propto r^{-\Gamma}$) to determine the impact of stellar ``youth''
% on SMBH feeding rates.
%
%\begin{figure}
%\includegraphics[width=\columnwidth]{NickPlot.eps}
%\caption{\label{NickPlot} SMBH feeding rates $\dot{M}=\eta(t) \times M_*(r_{\rm s})$, in arbitrary units.  The green, blue, and black curves are for galaxies with $\Gamma$ values of $0.1$, $0.5$, and $0.9$, respectively.  Solid curves represent impulsive-mode star formation, while dotted curves represent continuous-mode star formation.}
%\end{figure}
%
%%In Fig. \ref{NickPlot} we plot $\dot{M}$, in arbitrary units, as a function of time, for three different stellar density profiles $\Gamma = \{1.1, 1.5, 1.9\}$ (which are normalized to have the same mass at an influence radius $r_{\rm soi}=10~{\rm pc}$ around a $10^7M_\odot$ SMBH).  We parametrize the wind velocity as 
%\begin{equation}
%\frac{v_{\rm w}}{\rm km~s^{-1}}=600-400\tanh\left( \frac{t-10^{7.5}~{\rm yr}}{10^{7}~{\rm yr}}\right).
%\end{equation}
%The ``impulsive burst'' mode of star formation produces large ($\sim 10$) differences between the three $\dot{M}$ curves at early times, when $r_{\rm s}$ is small, but smaller ($\sim 3$) differences at late times, when $r_{\rm s}$ is large.  We also plot, as dotted curves, a simple model for the ``continuous'' mode of star formation, where mass loss is calculated as $\bar{\eta} = \int\eta(t)dt/t_{\rm h} \approx 4$ and an average energy injection in the wind is calculated as $\bar{v_{\rm w}^2}=\int\eta(t)v_{\rm w}^2(t)dt/\bar{\eta} \approx (800~{\rm km~s}^{-1})^2$.  Interestingly, the continuous mode of star formation produces small differences from late-time $\dot{M}$ seen in cusp galaxies with impulsive star formation; however, continuous mode star formation decreases late-time $\dot{M}$ by an order of magnitude relative to impulsive star formation in core galaxies.

@@

\begin{align}
&\rs=\frac{7}{2}\frac{G \Mbh}{n \vw^2}\\
&\simeq 6 \, \pc \left(\MbhNorm\right) \left(\vwOFH\right)^{-2}
\label{eq:stag_simple}
\end{align}

@@

\[
\dot{M}=
\begin{cases}
   1.5 \times 10^{-3}  \left(\Mbheight\right)^{1.9} \left(\vwOFH\right)^{-4} \Msun \, \pyear& \text{for core galaxies} \\
   3.3 \times 10^{-3} \left(\Mbheight\right)^{1.4} \left(\vwOFH\right)^{-2}  \Msun \, \pyear  & \text{for cusp galaxies} 
\end{cases}
\label{eq:mdot_analytic}
\]

@@

%% AG-Apparently even in the low heating case, it is still possible to
%% an outflow (from the analytic results above the gas density must be
%% quite steep in this case). However, at low heating we will probably
%% become subject to a cooling instablity.  AG-I believe that in the
%% case of core galaxy, the negative root of the quadratic should be
%% taken...

@@

The stagnation radius scales linearly with $\Mbh$ (see
Equation~\ref{eq:stag_simple}), while the $\rIa$ is roughly
constant. For small $\Mbh$, $\rs$ is well inside $\rIa$ and SNes will
not contribute to the $\vwO$.  Above $\sim10^8 \Msun$ SNes will
represent an important contribution to the overall heating.
% Thus, for $\Mbh=10^6 \Msun$, we may have $\vwO\simeq 100$ km/s with
% heating mostly coming from main sequence winds. On the other hand,
% for $\Mbh \gsim 10^8 \Msun$, we may have $\vwO\simeq 500$ km/s, with
% the main contribution coming from type Ia SNe. These two regimes are
% illustrated in Figure~\ref{fig:vw_eff}.
%\end{enumerate}

@@

% AG-I am not comfortable saying type Ia SNe's don't contribute to
% heating for smaller mass BHs. If the cooling time for the SNe is
% long then it would be wrong to ignore the SNe contribution. The
% difference would be that the SNe heating would just be less steady.
% AG-Thornton and http://arxiv.org/pdf/1402.6695v2.pdf...
\end{enumerate}

@@

Gas should radiatively cool as well--either from metal line cooling
for $T\lsim 2\times 10^7$ K, or through free-free emission for $T\gsim
2\times 10^7$ K.  However, if the cooling rate is much smaller than
the rate of energy injection then radiative cooling may be safely
neglected.

@@

For $\vwO$=1000 km/s, it is apparent from Figure~\ref{fig:cooling}, that rad
